---
title: "Progression Risk Report"
author: ""
date: "`r format(Sys.time(), '%d %B %Y')`"
header-includes:
  - \usepackage{comment}
output: 
  pdf_document: 
    fig_caption: yes
    fig_height: 4
    number_sections: yes
    toc: yes
params: 
  path: '.'
  info.file: NULL 
---



```{r setup, include=FALSE}
library(BarrettsProgressionRisk)
library(pander)
library(ggrepel)
library(gtable)
library(grid)

knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)


path = params$path
#path='.'
rawFile = grep('raw', list.files(path, pattern='txt', full.names=T), value=T, ignore.case=T)
corrFile = grep('corr|fitted', list.files(path, pattern='txt', full.names=T), value=T,ignore.case=T)

#pr = predictRisk(path)
#save(pr,file='example/pr.Rdata')
load(file='pr.Rdata')


samples = c( sampleNames(pr,T), sampleNames(pr,F) )

blacklist = read.table(system.file("extdata", "qDNAseq_blacklistedRegions.txt", package="BarrettsProgressionRisk"), sep = '\t', header=T)


demoFile = params$info.file
withD = ifelse(length(demoFile) == 1, T, F)

```

Sample files read in:

* Raw file: _`r rawFile`_
* Fitted file: _`r corrFile`_
`r if(!withD) {"\\begin{comment}"}`
* Pathology & p53 IHC reported: _`r demoFile`_
`r if(!withD) {"\\end{comment}"}`

Samples provided for analysis from QDNAseq files: `r paste(samples, collapse=', ')`. 

`r pander(head(segmentedValues(pr)[,c(1:8)]), caption='Example of the segmented data.', justify='left')`

For processing and predictions these samples are presumed to be from a single patient. If this is not true please provide separate directories for the raw files from each patient.

\pagebreak

# Processing pipeline

## Segment genomes

- Exclude `r nrow(blacklist)` regions known to be poorly covered/highly homologous
- Apply the piecewise constant fit algorithm (or multi-pcf) to data after exclusion
- Remove any segments with fewer than 67 supporting bases (ie. equivalent to 1Mb).
- Evaluate residuals filter samples with error rates greater than the cutoff of 0.015.

### Passed QC

The following samples passed QC post-segmentation.
`r pander(subset(sampleResiduals(pr), Pass, select=c('sample','varMAD_median')))` 


`r if (length(sampleNames(pr,F)) <= 0) {"\\begin{comment}"}`
### Failed QC

The following samples failed QC post-segmentation.
`r pander(subset(sampleResiduals(pr), !Pass, select=c('sample','varMAD_median')))` 
`r if (length(sampleNames(pr,F)) <= 0) {"\\end{comment}"}`

## Post Segmentation Plots

```{r, warning=F, echo=F, fig.height=length(pr$segmented$seg.plots)*2}
plotSegmentData(pr)
```
\pagebreak

# Predictions

Risks per-sample were assigned based on percentile calibrations of the probabilities across all samples used to train the model.  Probabilities that fall in the *blue* are *low* risk, *yellow* are *moderate* risk, and *red* are *high* risk.

For the samples provided the risks are:
`r pander(predictions(pr), caption='Risk predictions per sample.', justify='left')`

## _Which plot provides a user with more useful information??_

### Prediction Plot A

```{r, fig.height=5}
recommendations = rx(pr)

#pm = plotModelPredictions()
showPredictionCalibration() + 
  geom_point(data=predictions(pr), aes(x=Probability,y=0.5), color='grey39', shape=18, size=5, show.legend = F) + 
  geom_point(data=predictions(pr), aes(x=Probability,y=0.5,color=Risk), shape=18, size=4, show.legend = F) + 
  geom_text_repel(data=predictions(pr), aes(x=Probability,y=0.5,label=Sample,angle=45), show.legend = F )
```

### Prediction Plot B

```{r, fig.height=5}
# TODO add predictions(pr) as geom_points on the plot and label them.
pm = plotModelPredictions()
bp = ggplot_build(pm)

preds = predictions(pr)

preds = cbind(preds, do.call(rbind,lapply(preds$Probability, function(p) {
  subset(bp[['data']][[1]], p < xmax & p > xmin, select=c('y','fill'))
})))

for (i in 1:nrow(preds))
  pm = pm + 
  geom_point(data=preds[i,], aes(Probability,y,color=Risk), color='grey39', shape=18, size=5) +
  geom_point(data=preds[i,], aes(Probability,y,color=Risk), color=preds[i,'fill'], shape=18, size=3) + 
  geom_text_repel(data=preds[i,], aes(Probability,y,label=Sample),nudge_y=1)

pm
```


`r if(!withD) {"\\begin{comment}"}`
Per-sample p53 IHC and pathology are included using the provided `r demoFile`.  

`r if(!withD) {"\\end{comment}"}`

```{r echo=F, eval=withD}
demo = read.table(demoFile, sep='\t', header=T)
recommendations = rx(pr,demoFile)
```

\pagebreak

## Rx 

Recommendations per pair of samples, presuming these samples come from different timepoint/endoscopies. Each sample was predicted indpendently and recommendations are evaluated with the following criteria:

1. Immediate RFA: *HGD* or *IMC* diagnosis or more than one consecutive *high* risk predictions.
2. Recheck 6-12 months: One *high* risk prediction or an *aberrant p53 IHC*.
3. Recheck endoscopy 12-24 months: One or more *moderate* risk predictions.
4. Regular surveillance 3-5 years: Two or more consecutive *low* risk predictions.


If no p53 IHC or pathology information is provided the risks alone are used to provide recommendations.

### Recommendations per pair of samples
```{r}
fonts = c('bold.italic','italic','plain','plain' )
riskCols = RColorBrewer::brewer.pal(11, "RdYlBu")[seq(1,11, 3)]
tt3 <- gridExtra::ttheme_minimal(
  core=list(bg_params = list(fill=riskCols[recommendations$rule], col=NA),
            fg_params=list(fontface=fonts[recommendations$rule]), col='white'),
  colhead=list(fg_params=list(col="black", fontface=4L)))

table = gridExtra::tableGrob(format(recommendations[,c('Time 1', 'Time 2', 'Rx')], core.just='left'),rows=NULL,theme=tt3  )

table = gtable_add_grob(table,
        grobs = rectGrob(gp = gpar(fill = NA, lwd = 2)),
        t = 2, b = nrow(table), l = 1, r = ncol(table))

margin = theme(plot.margin = unit(c(0,0), "cm"))
gridExtra::grid.arrange(table,nrow=1,ncol=1)
```


